---
title: "Analysing Air Quality Data"
author: "Manodhayan"
date: "6 August 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Air Pollution

- Pollution is now a common place term, that our ears are attuned to. We hear about the various forms of pollution and read about it through the mass media.

- Air pollution can further be classified into two sections- Visible air pollution and invisible air pollution

- The Ozone layer considered crucial for the existence of the ecosystems on the planet is depleting due to increased pollution.


## Getting into the Data

```{r cars, echo = TRUE}
head(airquality)
```

## Ozone vs Wind
```{r plot1}
library(ggplot2)
data<-na.omit(airquality)
ggplot(data=data,aes(Wind,Ozone))+geom_smooth(method='lm')
```

## Ozone vs Temp
```{r Temp}
ggplot(data=data,aes(Temp,Ozone))+geom_smooth(method='lm')
```

## Ozone vs Solar.R
```{r Solar}
ggplot(data=data,aes(Solar.R,Ozone))+geom_smooth(method='lm')
```

## Linear Model Fit Results

```{r final,echo=FALSE}
fit<- lm(Ozone~.,data=data)
par(mfrow=c(2,2))
plot(fit)
```

## Thank You
