---
title: "Reproducible Research- Assignment"
author: "Manodhayan"
output:
  html_document:
    keep_md: yes
  pdf_document: 
    keep_tex: yes
---


##Synopsis
The purpose of this project was to practice:

* downloading, loading and preprocessing data
* imputing missing values
* interpreting data to answer research questions


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Read Data
```{r}
activity<-read.csv("activity.csv")
```
#Process Data
```{r}
# Transform the date attribute to an actual date format
activity$date <- as.POSIXct(activity$date, format="%Y-%m-%d")

# Compute the weekdays from the date attribute
activity <- data.frame(date=activity$date, 
                           weekday=tolower(weekdays(activity$date)), 
                           steps=activity$steps, 
                           interval=activity$interval)

# Compute the day type (weekend or weekday)
activity <- cbind(activity, 
                      daytype=ifelse(activity$weekday == "saturday" | 
                                     activity$weekday == "sunday", "weekend", 
                                     "weekday"))

# Create the final data.frame
activity_processed <- data.frame(date=activity$date, 
                       weekday=activity$weekday, 
                       daytype=activity$daytype, 
                       interval=activity$interval,
                       steps=activity$steps)

head(activity_processed)
```
# Histogram of Total numbers of steps taken each day
```{r}
# Compute the total number of steps each day
total_steps_per_day <- aggregate(activity_processed$steps, by=list(activity_processed$date), FUN=sum, na.rm=TRUE)

# Rename the attributes
names(total_steps_per_day) <- c("date", "total")
head(total_steps_per_day)
```
```{r echo=TRUE}
hist(total_steps_per_day$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="blue", 
     xlab="Total number of steps", 
     ylim=c(0, 20), 
     main="Histogram of the total number of steps taken each day")
```

#Mean and median number of steps taken each day
```{r}
mean(total_steps_per_day$total)
median(total_steps_per_day$total)

```
#Time Series plot of Number of steps taken per day
```{r echo=TRUE}
# Compute the means of steps accross all days for each interval
mean_data <- aggregate(activity_processed$steps, 
                       by=list(activity_processed$interval), 
                       FUN=mean, 
                       na.rm=TRUE)

# Rename the attributes
names(mean_data) <- c("interval", "mean")
head(mean_data)

# Compute the time series plot
plot(mean_data$interval, 
     mean_data$mean, 
     type="l", 
     col="blue", 
     lwd=2, 
     xlab="Interval [minutes]", 
     ylab="Average number of steps", 
     main="Time-series of the average number of steps per intervals")
```
#The 5-minute interval that, on average, contains the maximum number of steps
```{r}
# We find the position of the maximum mean
max_pos <- which(mean_data$mean == max(mean_data$mean))

# We lookup the value of interval at this position
max_interval <- mean_data[max_pos, 1]
print(max_interval)
```
#Code to describe and show a strategy for imputing missing data
```{r}
#Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA's)
NA_count <- sum(is.na(activity_processed$steps))
print(NA_count)

#Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
# Find the NA positions
na_pos <- which(is.na(activity_processed$steps))

# Create a vector of means
mean_vec <- rep(mean(activity_processed$steps, na.rm=TRUE), times=length(na_pos))

#Create a new dataset that is equal to the original dataset but with the missing data filled in
activity_processed[na_pos, "steps"] <- mean_vec
head(activity_processed)

```
#Histogram of the total number of steps taken each day after missing values are imputed
```{r}
sum_data <- aggregate(activity_processed$steps, by=list(activity_processed$date), FUN=sum)

# Rename the attributes
names(sum_data) <- c("date", "total")

# Compute the histogram of the total number of steps each day
hist(sum_data$total, 
     breaks=seq(from=0, to=25000, by=2500),
     col="blue", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Histogram of the total number of steps taken each day\n(NA replaced by mean value)")
mean(sum_data$total)
median(sum_data$total)

```
#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
head(activity_processed)
# Load the lattice graphical library
library(lattice)

# Compute the average number of steps taken, averaged across all daytype variable
mean_data <- aggregate(activity_processed$steps, 
                       by=list(activity_processed$daytype, 
                               activity_processed$weekday, activity_processed$interval), mean)

# Rename the attributes
names(mean_data) <- c("daytype", "weekday", "interval", "mean")
head(mean_data)
# Compute the time serie plot
xyplot(mean ~ interval | daytype, mean_data, 
       type="l", 
       lwd=1, 
       xlab="Interval", 
       ylab="Number of steps", 
       layout=c(1,2))
```
